---
title: "Taxonomic_composition"
author: 
- name: Mauricio Hernández
  affiliation: Laboratory of Biotic Interactions, Centro de Investigación en Ciencias
    Biológicas, Universidad Autónoma de Tlaxcala, Tlaxcala, Mexico
  email:  delmer.hernandez@unah.edu.hn
- name: Stephanie Hereira-Pacheco
  affiliation: Laboratory of Soil Ecology, CINVESTAV-IPN, Ciudad de México, Mexico
  email: shereirap@gmail.com
- name: Yendi E. Navarro-Noya
  affiliation: Laboratory of Biotic Interactions, Centro de Investigación en Ciencias
    Biológicas, Universidad Autónoma de Tlaxcala, Tlaxcala, Mexico
  email: yendiebenezer.navarro.n@uatx.mx
date: "`r format(Sys.time(), '%d - %m - %Y')`"
output: 
  html_document:
    theme: flatly
    highlight: tango
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
---

# I. PLOTS OF TAXONOMIC COMPOSITION

### Loading libraries

```{r, warning=FALSE, message=FALSE}
library(phyloseq)
library(ggplot2)
library(vegan)
library(picante)
library(edgeR)
library("RColorBrewer")
library(scales)
library(grid)
library(reshape2)
library(scales)
library(viridis)
library(hrbrthemes)
library(tidyverse)
library(VennDiagram)
```

```{r, warning=FALSE, message=FALSE, eval=FALSE}
#Installing edgeR

if (!require("BiocManager", quietly = TRUE))
install.packages("BiocManager")

BiocManager::install("edgeR")
```


## PLOT RELATIVE ABUNDANCE AT THE PHYLUM LEVEL

### Load data
```{r }
metadata <- read.csv(file = "../data/metadata.csv", header = TRUE, row.names = 1)
asv_table <- read.csv("../data/feature_table.csv", header = TRUE, row.names = 1)
taxonomy <- read.csv("../data/taxonomy.csv", header = TRUE, row.names = 1)
```

### Create phyloseq objects
```{r, warning=FALSE, message=FALSE,}
SAM <- sample_data(metadata)
TAX <- tax_table(as.matrix(taxonomy))
ASV <- otu_table(asv_table, taxa_are_rows=TRUE)  
physeq <- merge_phyloseq(ASV, TAX, SAM)
```

### Data visualizations
```{r}
sample_names(physeq)
rank_names(physeq)
sample_variables(physeq)

# Convert to relative abundances 
relative <- transform_sample_counts(physeq = physeq, 
                                    function(ASV) ASV / sum(ASV))
```


## RELATIVE ABUNDACE BY TAXONOMIC LEVEL (MEAN AND STANDARD DEVIATION)

### Calculations
```{r}
data <-
  physeq %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x)100* x / sum(x)) %>%
  psmelt() %>%
  as_tibble()

```


## PLOT RELATIVE ABUNDANCE AT THE PHYLUM LEVEL

### Create color palette
```{r}
paleta <- c(brewer.pal(12, "Paired")[1:12], "gray", "orange")
print(paleta)
```

### Species + Season

```{r}
Rel_Abun_Phyl_Season <- plot_bar(physeq = relative, "Sample", fill = "Phylum") + 
  facet_grid(~Species_Scel + Season, space = "free", scales = "free") +
  labs(y="Relative abundance") +
  geom_bar(stat = "identity", pisition = "stack", res = 300) +
  scale_fill_manual(values = paleta) +
  theme(strip.text.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold")) +
  theme(strip.text.x = element_text(face = "italic")) +
  theme(legend.text = element_text(face = "italic")) +
  theme(strip.background = element_rect(colour = "black", fill = "white")) +
  theme(text = element_text(size = 10))
print(Rel_Abun_Phyl_Season)
ggsave("../figures/Rel_Abun_Phyl_Season.jpeg", width=15, height=4.5, dpi=300)
```


## PLOT RELATIVE ABUNDANCE AT THE GENUS LEVEL


### Load data
```{r}
metadata <- read.csv(file = "../data/metadata.csv", check.names = F)
asv_table <- read.csv("../data/feature_table.csv", check.names = F)
taxonomy <- read.csv("../data/taxonomy_03.csv", check.names = F) %>% unite(
  taxa, Kingdom:Species, remove = F, sep = ";")
asvtable_metadata <- asv_table %>% 
  inner_join(taxonomy)
```


## Relative abundance at phylum level (Sceloporus species)

```{r}
Genus_01 <- asvtable_metadata %>% 
  group_by(Genus) %>% # But also for other taxonomic levels
  summarise_if(is.numeric, sum)
Genus_01 <- Genus_01 %>% 
  column_to_rownames(var = "Genus")
Genus.ra <- t(t(Genus_01)/colSums(Genus_01)*100)

rowMeans(Genus.ra) %>% 
  as.data.frame() %>% arrange(desc(.))
apply(Genus.ra,1,sd)

Genus <- Genus.ra %>% t() %>% 
  as.data.frame()%>% 
  rownames_to_column(var = "SampleID") %>% 
  inner_join(metadata)

Mean_Genus <- Genus %>% 
  group_by(Species_Scel) %>% 
  summarise_if(is.numeric, mean)
SD_Genus <- Genus %>% 
  group_by(Species_Scel) %>% 
  summarise_if(is.numeric, sd)
aggregate(Genus[ ,2:13], list(Genus$Species_Scel), mean)
aggregate(Genus[ ,2:13], list(Genus$Species_Scel), sd)
#write.table(Prom_Genus, file="./Genus_Pro_ST.txt", sep = "\t")
```

## Filter genus taxonomic level
```{r}
metadata <- read.csv(file = "../data/metadata.csv", header = TRUE, row.names = 1) 
asv_table <- read.csv(file = "../data/feature_table.csv", check.names = F) 
taxonomy <-  read.csv("../data/taxonomy_03.csv", check.names = F) %>% mutate_at(
  c("Genus"), str_replace,"g__", "")
```

```{r}
lista <- rowMeans(Genus.ra) %>% as.data.frame() %>% 
  arrange(desc(.)) %>% 
  slice_head(n=14) %>% 
  rownames_to_column(var = "Genus") %>% 
  filter(!Genus =="g__") %>% 
  mutate_at(c("Genus"), str_replace,"g__", "")
list <- lista$Genus
lista01 <- read.csv(file = "../data/lista.csv", check.names = F)
list02 <- lista01$Genus
#write.table(lista, file="../data/lista.txt", sep = "\t")
```

```{r}
taxonomy_filter <- taxonomy %>% 
  filter(Genus %in% list02)
taxonomy %>% inner_join(lista)
taxonomy_1 <- taxonomy_filter %>% 
  inner_join(asv_table, by =c(
  "OTUID"="OTUID")) %>% dplyr::select(1:8)
```

### Load data again
```{r}
asv_table_1 <- read.csv(file = "../data/feature_table.csv", header = TRUE,
                        row.names = 1) %>% 
  rownames_to_column(var = "OTUID") %>% 
  inner_join(taxonomy_1, by = "OTUID") %>% 
  dplyr::select(-97:-103) %>% 
  column_to_rownames(var = "OTUID")
taxo <- taxonomy_1 %>% column_to_rownames(var = "OTUID")
```

### Create phyloseq objects
```{r}
SAM <- sample_data(metadata)
TAX <- tax_table(as.matrix(taxo)) 
OTU <- otu_table(asv_table_1, taxa_are_rows=TRUE)  
physeq <- merge_phyloseq(OTU, TAX, SAM)
```

### Convert to relative abundance
```{r}
relative  = transform_sample_counts(physeq = physeq, function(OTU) OTU / sum(OTU))
```

### Color palette
```{r}
paleta <- c(brewer.pal(12, "Paired")[1:12], brewer.pal(8, "Dark2")[1:8])
print(paleta)
```

## Plot by Species and Season 
```{r}
Rel_Abun_Genus_Season <- plot_bar(physeq = relative, "Sample", fill = "Genus") + 
  facet_grid(~Species_Scel + Season, space = "free", scales = "free") +
  labs(y="Relative abundance") +
  geom_bar(stat = "identity", pisition = "stack", res = 300) +
  scale_fill_manual(values = paleta) +
  theme(strip.text.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold")) +
  theme(strip.text.x = element_text(face = "italic")) +
  theme(legend.text = element_text(face = "italic")) +
  theme(strip.background = element_rect(colour = "black", fill = "white")) +
  theme(text = element_text(size = 10))
print(Rel_Abun_Genus_Season)
ggsave("../figures/Rel_Abun_Genus_Season.jpeg", width=15.0, height=6.5, dpi=400)
```



## Core microbiota (50%) 

```{r}
aeneus_50 <- read.delim("../data/core_50_aeneus.tsv", sep = "\t", check.names = F) %>%
  rownames_to_column(var = "ids")

bicanthalis_50 <- read.delim("../data/core_50_bicanthalis.tsv", sep = "\t", check.names = F) %>%
  rownames_to_column(var = "ids")

grammicus_50 <- read.delim("../data/core_50_grammicus.tsv", sep = "\t", check.names = F) %>%
  rownames_to_column(var = "ids")

spinosus_50 <- read.delim("../data/core_50_spinosus.tsv", sep = "\t", check.names = F) %>%
  rownames_to_column(var = "ids") 
```


### Plotting the Venn Diagramm

```{r}
venn.plot_50 <- venn.diagram(
  x = list(S_grammicus = grammicus_50$OTUID,
           S_spinosus = spinosus_50$OTUID,
           S_aeneus = aeneus_50$OTUID,
           S_bicanthalis = bicanthalis_50$OTUID),
  category.names = c(
    expression(bold("S. grammicus")),
    expression(bold("S. spinosus")),
    expression(bold("S. aeneus")),
    expression(bold("S. bicanthalis"))),
  filename = "../figures/viendo_50.jpg",
  output = TRUE,
  height = 3000,
  width = 6000,
  resolution = 300,
  compression = "lzw",
  units = "px",
  lwd = 6,
  lty = "blank",
  fill = c("#56B4E9", "#009E73", "#999999", "#E69F00"),
  cex = 1.5,
  #fontface = "bold",
  fontface = "italic",
  fontfamily = "sans",
  cat.cex = 2,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.pos = c(-10, 25, -0.020, -0.050),
  cat.dist = c(0.3, 0.35, 0.099, 0.099),
  cat.fontfamily = "sans")

```

