---
title: "Alpha_div_tax"
author: 
- name: Mauricio Hernández
  affiliation: Laboratory of Biotic Interactions, Centro de Investigación en Ciencias
    Biológicas, Universidad Autónoma de Tlaxcala, Tlaxcala, Mexico
  email:  delmer.hernandez@unah.edu.hn
- name: Stephanie Hereira-Pacheco
  affiliation: Laboratory of Biotic Interactions, Centro de Investigación en Ciencias
    Biológicas, Universidad Autónoma de Tlaxcala, Tlaxcala, Mexico
  email: shereirap@gmail.com
- name: Sergio Ancona
  affiliation: Departamento de Ecología Evolutiva, Instituto de Ecología, Universidad Nacional Autónoma de México
  email:  sergio.ancona@iecologia.unam.mx
- name: Yendi E. Navarro-Noya
  affiliation: Laboratory of Biotic Interactions, Centro de Investigación en Ciencias
    Biológicas, Universidad Autónoma de Tlaxcala, Tlaxcala, Mexico
  email: yendiebenezer.navarro.n@uatx.mx
date: "`r format(Sys.time(), '%d - %m - %Y')`"
output: 
  html_document:
    theme: flatly
    highlight: tango
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# II. ALPHA DIVERSITY ANALYSES

## Loading libraries
```{r, warning=FALSE, message=FALSE}
library(hilldiv)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(tidyverse)
library(lme4)
library(car)
library(arm)
library(predictmeans)
library(AER)
library(visreg)
library(FSA)
library(rstatix)
library(sjPlot)
library(effects)
library(cowplot)
```

## Calculating Hill numbers with hilldiv R package

```{r, eval=FALSE}
otu_table <- read.csv("feature_table.csv", header = TRUE, row.names = 1)
q0 <- hill_div(otu_table, qvalue = 0)
q1 <- hill_div(otu_table, qvalue = 1)
q2 <- hill_div(otu_table, qvalue = 2)
q012 <- cbind(q0, q1, q2)
write.table(q012, file="../data/q012_hilldiv.txt", sep = "\t")
```

## Preparing the data

### Load files
```{r, highlight=TRUE}
richness_q012 <- read.csv("../data/Hill_numbers_q012.csv", header = TRUE) %>% 
  dplyr::select(SampleID, q0, q1, q2)
metadata <- read.csv("../data/metadata.csv", header = TRUE, check.names = F)
Micro_div <- richness_q012 %>% 
  inner_join(metadata, by = c("SampleID"="SampleID"))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, out.width=8}
DT::datatable(
  Micro_div,
  fillContainer = FALSE, 
  filter="top")
```

### Declare factors
```{r}
SPECIES <- as.factor(Micro_div$Species_Scel) # Four Sceloporus lizard species
SEASON <- as.factor(Micro_div$Season) # Dry and Rainy
SEX <- as.factor(Micro_div$Sex) # Male and Female
```


### Standardize continous independent variables
```{r}
SVL <- rescale(Micro_div$SVL, binary.inputs = "center") # Snout–vent length measured in mm.
ELEVATION <- rescale(Micro_div$Elevation, binary.inputs = "center") # Taken as m a.s.l. 
SEQDEPTH <- rescale(Micro_div$SeqDepth, binary.inputs = "center")
```

## Evaluate whether sex variable influence on gut microbiota diversity (q=1)

```{r}
# Paired test (Wilcoxon test, q1) 

# Sceloporus aeneus
aeneus_fem_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus aeneus" & 
                            Sex == "Female", q1, drop = TRUE)
aeneus_male_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus aeneus" & 
                             Sex == "Male", q1, drop = TRUE)
aeneus_FemMale_GutMic <- wilcox.test(x= aeneus_fem_GutMic, y= aeneus_male_GutMic)
aeneus_FemMale_GutMic

# Sceloporus bicanthalis
bica_fem_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus bicanthalis" & 
                          Sex == "Female", q1, drop = TRUE)
bica_male_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus bicanthalis" & 
                           Sex == "Male", q1, drop = TRUE)
bica_FemMale_GutMic <- wilcox.test(x= bica_fem_GutMic, y= bica_male_GutMic)
bica_FemMale_GutMic

# Sceloporus grammicus
gram_fem_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus grammicus" & 
                          Sex == "Female", q1, drop = TRUE)
gram_male_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus grammicus" & 
                           Sex == "Male", q1, drop = TRUE)
gram_FemMale_GutMic <- wilcox.test(x= gram_fem_GutMic, y= gram_male_GutMic)
gram_FemMale_GutMic

# Sceloporus spinosus
spi_fem_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus spinosus" & 
                         Sex == "Female", q1, drop = TRUE)
spi_male_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus spinosus" & 
                          Sex == "Male", q1, drop = TRUE)
spi_FemMale_GutMic <- wilcox.test(x= spi_fem_GutMic, y= spi_male_GutMic)
spi_FemMale_GutMic

```


## Differences in  snout vent length (SVL) among lizard species
```{r}
set.seed(1234)
Data_SVL <- subset(Micro_div, select = c(Species_Scel, SVL))
kruskal.t <- Data_SVL %>% kruskal_test(SVL ~ Species_Scel)
kruskal.t
Post_hoc  <- Data_SVL %>% dunn_test(SVL ~ Species_Scel, p.adjust.method = "bonferroni") 
Post_hoc

```



### Visualization: box plots with p-values
```{r}
Post_hoc <- Post_hoc %>% 
  add_xy_position(x = "Species_Scel")
ggboxplot(Data_SVL, x = "Species_Scel", y = "SVL", fill = "Species_Scel") +
  xlab(element_blank())+
  scale_fill_manual(values = c("#56B4E9","#009E73", "#999999","#E69F00"))+
      theme_classic() +
   theme(legend.position = "right",
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank()) +
  theme(legend.text = element_text(face = "italic"))+
  stat_pvalue_manual(Post_hoc, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(kruskal.t, detailed = TRUE),
    caption = get_pwc_label(Post_hoc))
```


## Generalized Linear Model with quasi-Poisson distribution (q = 1) 

```{r}
M1 <- glm(q1 ~ SEASON*SPECIES+ELEVATION+SVL+SEQDEPTH,
           family = quasipoisson(link = "log"),
           data = Micro_div)
summary(M1)
plot(M1)

M2 <- update(M1,. ~ . -SPECIES:SEASON)
anova(M1, M2, test = "F")
summary(M2)

M3 <- update(M2,. ~ . -SVL)
anova(M2, M3, test = "F")
summary(M3)

M4 <- update(M3,. ~ . -ELEVATION) 
anova(M3, M4, test = "F")
summary(M4)
confint(M4)

M5 <- update(M4,. ~ . -SEQDEPTH) 
anova(M4, M5, test="F")
summary(M5)

# Best Model
M6 <- glm(q1 ~ SEASON*SPECIES,
          family = quasipoisson(link = "log"),
          data = Micro_div)
anova(M6, test="F")
summary(M6)
confint(M6)
coef(M6)
```



## Residual plot of the best model
### Plot model assumption

```{r}
plot_model(M6, type = "eff", terms = "SPECIES")
plot_model(M6, show.values = TRUE, value.offset = .3, width = 0.1,
           vline.color = "#E69F00")

residuals_q1 <- plot_model(M6, type = "eff", terms = c("SEASON","SPECIES"),
                           colors = c("#999999", "#E69F00", "#56B4E9", 
                                      "#009E73", "#F0E442")) +
  theme_classic() +
  theme(legend.text = element_text(face = "italic")) +
  theme(axis.text.x = element_text(size = 11))
residuals_q1 
#ggsave("../figures/residuals_q1.jpeg", width = 5.0, height = 3.0, dpi = 300)
```


### Contrasts (dry vs rainy), CI95%, p values
#### Tukey Test

```{r}
library(emmeans)
emmeans(M6, specs = ~ SPECIES*SEASON) %>%
  contrast() %>%
  as.data.frame()
emm <- emmeans(M6, spec = ~ SPECIES*SEASON, 
               type = "response")
print(emm)
X <- contrast(emm, method = "pairwise")
confint(X)

M1.emm.s <- emmeans(M6, specs = ~ SPECIES*SEASON)
pairs(M1.emm.s, adjust = "tukey", infer=c(TRUE,TRUE))
pwpm(M1.emm.s)
summary(M1.emm.s, type="response")
plot(M1.emm.s, comparisons=TRUE, type="response")

```


## Evaluate whether sex variable influence on gut microbiota diversity (q=2)

```{r}
# Paired test (Wilcoxon test, q2) 

# Sceloporus aeneus
aeneus_fem_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus aeneus" & 
                            Sex == "Female", q2, drop = TRUE)
aeneus_male_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus aeneus" & 
                             Sex == "Male", q2, drop = TRUE)
aeneus_FemMale_GutMic <- wilcox.test(x= aeneus_fem_GutMic, y= aeneus_male_GutMic)
aeneus_FemMale_GutMic

# Sceloporus bicanthalis
bica_fem_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus bicanthalis" & 
                          Sex == "Female", q2, drop = TRUE)
bica_male_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus bicanthalis" & 
                           Sex == "Male", q2, drop = TRUE)
bica_FemMale_GutMic <- wilcox.test(x= bica_fem_GutMic, y= bica_male_GutMic)
bica_FemMale_GutMic

# Sceloporus grammicus
gram_fem_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus grammicus" & 
                          Sex == "Female", q2, drop = TRUE)
gram_male_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus grammicus" & 
                           Sex == "Male", q2, drop = TRUE)
gram_FemMale_GutMic <- wilcox.test(x= gram_fem_GutMic, y= gram_male_GutMic)
gram_FemMale_GutMic

# Sceloporus spinosus
spi_fem_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus spinosus" & 
                         Sex == "Female", q2, drop = TRUE)
spi_male_GutMic <- subset(Micro_div, Species_Scel == "Sceloporus spinosus" & 
                          Sex == "Male", q2, drop = TRUE)
spi_FemMale_GutMic <- wilcox.test(x= spi_fem_GutMic, y= spi_male_GutMic)
spi_FemMale_GutMic
```


## Generalized Linear Model with quasi-Poisson distribution (q = 2)

```{r}
MODEL1 <- glm(q2 ~ SEASON*SPECIES+ELEVATION+SVL+SEQDEPTH,
            family = quasipoisson(link = "log"),
            data = Micro_div)
summary(MODEL1)
plot(MODEL1)

MODEL2 <- update(MODEL1,. ~ . -SPECIES:SEASON)
anova(MODEL1, MODEL2, test = "F")
summary(MODEL2)

MODEL3 <- update(MODEL2,. ~ . -SVL)
anova(MODEL2, MODEL3, test="F")
summary(MODEL3)

MODEL4 <- update(MODEL3,. ~ . -SEQDEPTH)
anova(MODEL3, MODEL4, test="F")
summary(MODEL4)

MODEL5 <- update(MODEL4,. ~ . -ELEVATION) # Best model
anova(MODEL4, MODEL5, test="F")
summary(MODEL5)

# Best Model
MODEL6 <- glm(q2 ~ SEASON*SPECIES,
              family = quasipoisson(link = "log"),
              data = Micro_div)
summary(MODEL6)
coef(MODEL6)
confint(MODEL6)

```


### Residual plot of the best model
#### Plot model assumption
```{r}

plot_model(MODEL6, type = "eff", terms = "SPECIES")
plot_model(MODEL6, show.values = TRUE, value.offset = .3, width = 0.1,
           vline.color = "#E69F00")

residuals_q2 <- plot_model(MODEL6, type = "eff", terms = c("SEASON","SPECIES"),
                           colors = c("#999999", "#E69F00", "#56B4E9", 
                                      "#009E73", "#F0E442")) +
  theme_classic() +
  theme(legend.text = element_text(face = "italic")) +
  theme(axis.text.x = element_text(size = 11))
residuals_q2 
#ggsave("../figures/residuals_q2.jpeg", width = 5.0, height = 3.0, dpi = 300)

```


#### Contrasts (dry vs rainy), CI95%, p values


```{r}
## Tukey Test

emmeans(MODEL6, specs = ~ SPECIES*SEASON) %>%
  contrast() %>%
  as.data.frame()

emm_q2 <- emmeans(MODEL6, spec = ~ SPECIES*SEASON, 
                  type = "response")
print(emm_q2)
X_q2 <- contrast(emm_q2, method = "pairwise")
confint(X_q2)

M1.emm.s_q2 <- emmeans(MODEL6, specs = ~ SPECIES*SEASON)
pairs(M1.emm.s_q2, adjust = "tukey", infer=c(TRUE,TRUE))
pairs(M1.emm.s_q2, adjust = "tukey")
pwpm(M1.emm.s_q2)
summary(M1.emm.s_q2, type="response")
plot(M1.emm.s_q2, comparisons=TRUE, type="response")

```


## Relation between alpha taxonomic microbiota and diet

```{r}
Micro_div$Ind <- substr(Micro_div$SampleID, 8, nchar(Micro_div$SampleID) - 0)

diet_phylo_div <- read.csv("../data/diet_Hill_numbers_q012.csv", header = TRUE) 
# Generar una nueva columna con la información de otra columna, excluyendo los últimos 4 caracteres
diet_phylo_div$Ind <- substr(diet_phylo_div$SampleID, 8, nchar(diet_phylo_div$SampleID) - 4)
#names(diet_phylo_div)[names(diet_phylo_div) %in% c("q0", "q1", "q2")] <- c("d.p.0", "d.p.1", #"d.p.2")
divs <- Micro_div %>% 
  inner_join(diet_phylo_div, by = c("Ind"="Ind"))

#write.table(divs, file="../data/full_divs.txt", sep = "\t")

##Plot

taxq1_tq0 <- ggscatter(divs, x = "tq0", y = "q1", color = "Species_Scel",
   palette = c("#999999", "#E69F00", "#56B4E9", "#009E73"), xlab= "Taxonomic richness of diet (number of taxonomic groups)", ylab="Gut bacterial diversity (effective number of frequent ASVs)",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")
   )+
  theme(legend.title = element_blank(), legend.position = "none")


taxq2_tq0 <- ggscatter(divs, x = "tq0", y = "q2", color = "Species_Scel",
   palette = c("#999999", "#E69F00", "#56B4E9", "#009E73"), 
   xlab= "Taxonomic richness of diet (number of taxonomic groups)", ylab="Gut bacterial diversity (effective number of dominant ASVs)",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")
   )+
  theme(legend.title = element_blank(), legend.position = "none")


phyq1_pq0 <- ggscatter(divs, x = "pq0", y = "q1", color = "Species_Scel",
   palette = c("#999999", "#E69F00", "#56B4E9", "#009E73") , 
   xlab= " Phylogenetic richness of diet (number of phylogenetic lineages)", ylab="Gut bacterial diversity (effective number of frequent ASVs)",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")
   )+
  theme(legend.title = element_blank(), legend.position = "none")


phyq2_pq0 <- ggscatter(divs, x = "pq0", y = "q2", color = "Species_Scel",
   palette = c("#999999", "#E69F00", "#56B4E9", "#009E73"), 
   xlab= "Phylogenetic richness of diet (number of phylogenetic lineages)", 
   ylab="Gut bacterial diversity (effective number of dominant ASVs)",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")
   )+
  theme(legend.title = element_blank(), legend.position = "none")


taxq1_tq0.sp <- ggscatter(divs, x = "tq0", y = "q1", color = "Species_Scel",
   palette = c("#999999", "#E69F00", "#56B4E9", "#009E73"), 
   xlab= "Taxonomic richness of diet (number of taxonomic groups)", 
   ylab="Gut bacterial diversity (effective number of frequent ASVs)",
   facet.by = "Species_Scel",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", 
                         label.x = 3, label.sep = "\n")
   )+
  theme(legend.position = "none",
        strip.text = element_text(face = "italic"))



taxq2_tq0.sp <- ggscatter(divs, x = "tq0", y = "q2", color = "Species_Scel",
   palette = c("#999999", "#E69F00", "#56B4E9", "#009E73"), 
   xlab= "Taxonomic richness of diet (number of taxonomic groups)", ylab="Gut bacterial diversity (effective number of dominant ASVs)",
   facet.by = "Species_Scel" ,
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")
   )+
  theme(legend.position = "none",
        strip.text = element_text(face = "italic"))


phyq1_pq0.sp <- ggscatter(divs, x = "pq0", y = "q1", color = "Species_Scel",
   palette = c("#999999", "#E69F00", "#56B4E9", "#009E73") , 
   xlab= " Phylogenetic richness of diet (number of phylogenetic lineages)", ylab="Gut bacterial diversity (effective number of frequent ASVs)",
   facet.by = "Species_Scel",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")
   )+
  theme(legend.position = "none",
        strip.text = element_text(face = "italic"))


phyq2_pq0.sp <- ggscatter(divs, x = "pq0", y = "q2", color = "Species_Scel",
   palette = c("#999999", "#E69F00", "#56B4E9", "#009E73"), 
   xlab= "Phylogenetic richness of diet (number of phylogenetic lineages)", ylab="Gut bacterial diversity (effective number of dominant ASVs)", 
   facet.by = "Species_Scel" ,
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")
   )+
  theme(legend.position = "none",
        strip.text = element_text(face = "italic"))

taxq1_tq0
taxq1_tq0.sp
taxq2_tq0
taxq2_tq0.sp
phyq1_pq0
phyq1_pq0.sp
phyq2_pq0
phyq2_pq0.sp

diet.vs.micro1<- plot_grid(taxq1_tq0, taxq1_tq0.sp, 
                          labels = "AUTO", ncol = 1)
#ggsave("../figures/dietvsmicro1.jpeg", width=8, height=11, dpi=300)
diet.vs.micro2<- plot_grid(taxq2_tq0, taxq2_tq0.sp, 
                          labels = c("E","F"), ncol = 1)
#ggsave("../figures/dietvsmicro2.jpeg", width=8, height=11, dpi=300)
diet.vs.micro3<- plot_grid(phyq1_pq0, phyq1_pq0.sp, 
                          labels = c("C", "D"), ncol = 1)
#ggsave("../figures/dietvsmicro3.jpeg", width=8, height=11, dpi=300)
diet.vs.micro4<- plot_grid(phyq2_pq0, phyq2_pq0.sp, 
                          labels = c("G", "H"), ncol = 1)
#ggsave("../figures/dietvsmicro4.jpeg", width=8, height=11, dpi=300)

```

---

## Effect of diet on gut microbiota

### loading files
```{r}
richness_microb <- read.csv("../data/Hill_numbers_q012.csv", header = TRUE) %>% 
  dplyr::select(SampleID, q0, q1, q2)
metadata <- read.csv("../data/Metadata_model_DietMic.csv", header = TRUE, check.names = F)
Microb_data <- richness_microb %>% inner_join(metadata, by = c("SampleID"="SampleID"))
```


### Check distribution of the dependent variable (diversity q=0, Hill numbers)
```{r}
shapiro.test(Microb_data$q0) # Data are not normal
shapiro.test(Microb_data$q1) # Normal data

hist(Microb_data$q0, col = "darkgreen") # q0
hist(log(Microb_data$q0), col = "orange") # log transformation
hist(Microb_data$q1, col = "gray") # q1
hist(Microb_data$q2, col = "blue") # q2

bartlett.test(q0 ~ Species_Scel, data = Microb_data) 
leveneTest(q0 ~ Species_Scel, data = Microb_data)
```



### Plot variances among lizard species (q0)
```{r}
boxplot(q0 ~ Species_Scel, data = Microb_data,
        main = "Differences in variance among lizard species",
        xlab = "Species",
        ylab = " Hill numbers (q0)",
        col = "#4682B433",
        border = "black")

```


###Preparing the data

```{r}
#Declare as factors
SPECIES <- as.factor(Microb_data$Species_Scel) # Four Sceloporus lizard species
SEASON <- as.factor(Microb_data$Season) # Dry and Rainy
SEX <- as.factor(Microb_data$Sex) # Male and Female

#Standardize continous independent variables
SVL <- rescale(Microb_data$SVL, binary.inputs = "center") # Snout–vent length measured in mm.
ELEVATION <- rescale(Microb_data$Elevation, binary.inputs = "center") # Taken as m a.s.l. 
SEQDEPTH <- rescale(Microb_data$SeqDepth, binary.inputs = "center")
```




### Linear model with Poisson distribution (taxonomic q = 0) 

#### With Microbial Diversity (q1) 
```{r}
## UPDATE FUNCTION (RUN ANOVA)
# 68 samples (diet / microbiota)
m1 <- glm(q1 ~ Diet*SPECIES+SEASON,
          family = quasipoisson(link = "log"),
          data = Microb_data)
summary(m1)
plot(m1)

m2 <- update(m1,. ~ . -Diet:SPECIES)
anova(m1, m2, test="F")
summary(m2)

m3 <- update(m2,. ~ . -Diet)
anova(m2, m3, test="F")
summary(m3) # Best Model

m4 <- update(m3,. ~ . -SPECIES)
anova(m3, m4, test="F")
summary(m4)

m5 <- update(m3,. ~ . -SEASON)
anova(m3, m5, test="F")
summary(m5)

# Plot model assumption
plot_model(m3, type = "eff", terms = "SPECIES",
           colors = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  theme(legend.text = element_text(face = "italic"))

plot_model(m3, type = "eff", terms = c("SPECIES", "SEASON"),
           colors = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  theme_classic() +
  theme(legend.text = element_text(face = "italic")) +
  theme(axis.text.x = element_text(size = 11))
```


#### With Microbial Diversity (q2) 

```{r}
model1 <- glm(q2 ~ Diet*SPECIES+SEASON,
              family = quasipoisson(link = "log"),
              data = Microb_data)
summary(model1)
plot(model1)

model2 <- update(model1,. ~ . -Diet:SPECIES)
anova(model1, model2, test="F")
summary(model2)

model3 <- update(model2,. ~ . -Diet)
anova(model2, model3, test="F")
summary(model3)

model4 <- update(model3,. ~ . -SEASON)
anova(model3, model4, test="F") # Best model
summary(model4)

model5 <- update(model4,. ~ . -SPECIES)
anova(model4, model5, test="F")
summary(model5)

# Plot model assumption
plot_model(model3, type = "eff", terms = "SPECIES",
           colors = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  theme(legend.text = element_text(face = "italic"))
```

### Linear models with Poisson  distribution (Phylogenetic q = 0) 

#### With Microbial Diversity (q1)
```{r}

divs$phyDiet<- divs$pq0
#Declare as factors
SPECIES <- as.factor(divs$Species_Scel) # Four Sceloporus lizard species
SEASON <- as.factor(divs$Season) # Dry and Rainy
SEX <- as.factor(divs$Sex) # Male and Female

#Standardize continous independent variables
SVL <- rescale(divs$SVL, binary.inputs = "center") # Snout–vent length measured in mm.
ELEVATION <- rescale(divs$Elevation, binary.inputs = "center") # Taken as m a.s.l. 
SEQDEPTH <- rescale(divs$SeqDepth, binary.inputs = "center")

## UPDATE FUNCTION (RUN ANOVA)
# 68 samples (diet / microbiota)
m1 <- glm(q1 ~ phyDiet*SPECIES+SEASON,
          family = quasipoisson(link = "log"),
          data = divs)
summary(m1)
plot(m1)

m2 <- update(m1,. ~ . -phyDiet:SPECIES)
anova(m1, m2, test="F")
summary(m2)

m3 <- update(m2,. ~ . -phyDiet)
anova(m2, m3, test="F")
summary(m3) # Best Model

m4 <- update(m3,. ~ . -SPECIES)
anova(m3, m4, test="F")
summary(m4)

m5 <- update(m3,. ~ . -SEASON)
anova(m3, m5, test="F")
summary(m5)

# Plot model assumption
plot_model(m3, type = "eff", terms = "SPECIES",
           colors = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  theme(legend.text = element_text(face = "italic"))

plot_model(m3, type = "eff", terms = c("SPECIES", "SEASON"),
           colors = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  theme_classic() +
  theme(legend.text = element_text(face = "italic")) +
  theme(axis.text.x = element_text(size = 11))
```


#### With Microbial Diversity (q2) 

```{r}
model1 <- glm(q2 ~ phyDiet*SPECIES+SEASON,
              family = quasipoisson(link = "log"),
              data = divs)
summary(model1)
plot(model1)

model2 <- update(model1,. ~ . -phyDiet:SPECIES)
anova(model1, model2, test="F")
summary(model2)

model3 <- update(model2,. ~ . -phyDiet)
anova(model2, model3, test="F")
summary(model3)

model4 <- update(model3,. ~ . -SEASON)
anova(model3, model4, test="F") # Best model
summary(model4)

model5 <- update(model4,. ~ . -SPECIES)
anova(model4, model5, test="F")
summary(model5)

# Plot model assumption
plot_model(model3, type = "eff", terms = "SPECIES",
           colors = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  theme(legend.text = element_text(face = "italic"))
```
