---
title: "Beta_diversity_analysis"
author: 
- name: Mauricio Hernández
  affiliation: Laboratory of Biotic Interactions, Centro de Investigación en Ciencias
    Biológicas, Universidad Autónoma de Tlaxcala, Tlaxcala, Mexico
  email:  delmer.hernandez@unah.edu.hn
- name: Yendi E. Navarro-Noya
  affiliation: Laboratory of Biotic Interactions, Centro de Investigación en Ciencias
    Biológicas, Universidad Autónoma de Tlaxcala, Tlaxcala, Mexico
  email: yendiebenezer.navarro.n@uatx.mx
date: "`r format(Sys.time(), '%d - %m - %Y')`"
output: 
  html_document:
    theme: flatly
    highlight: tango
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
---

# III. BETA DIVERSITY ANALYSES

### Install hilldiv2
```{r, eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("anttonalberdi/hilldiv2")
```

##  Loading libraries
```{r, warning=FALSE, message=FALSE}
library(hilldiv2)
library(tidyverse)
library(vegan)
library(phytools)
library(pairwiseAdonis)
library(reshape2)
library(ggpubr)
library(rstatix)
```

```{r}
# Load files
otutable <- read.csv("../data/feature_table.csv", row.names = 1)
Metadata <- read.csv("../data/metadata_beta.csv")
```


## Calculate beta diversity

```{r}
hill_pair_dis <- hillpair(data = otutable, q=c(1,2))
names(hill_pair_dis)
head(hill_pair_dis$q1S)
```
- S (Jaccard-type turnover): it quantifies the normalised species turnover rate in a sample relative to the total pool of samples (i.e., gamma diversity).
- V (Sørensen-type turnover): it quantifies the normalised species turnover rate in a sample relative to one sample (i.e., alpha diversity).
- U (Jaccard-type overlap-complement): it quantifies the proportion of non-shared species in the total pool of samples. Therefore, this metric quantifies dissimilarity from the perspective of the pool of samples.
- C (Sørensen-type overlap-complement): it quantifies the effective average proportion of non-shared OTUs/ASVs/MAGs in samples. Therefore, this metric quantifies dissimilarity from the perspective of a single sample.

## Taxonomic beta diversity  at order q1 #

```{r, warning=FALSE, message=FALSE, cache.comments=FALSE, include=FALSE}
hill_pair_dis_nmds <- hill_pair_dis$q1S %>% #select the distance object based on dissimilarity metric S
  metaMDS(.,trymax = 500, k = 2, verbosity = FALSE) %>%
  vegan:::scores.metaMDS() %>%
  as_tibble(., rownames = "sample")

hill_pair_dis_nmds <- hill_pair_dis_nmds %>%
  left_join(Metadata, by = join_by(sample == SampleID)) %>%
  group_by(Species, Season) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup()
```

```{r, eval=FALSE}
hill_pair_dis_nmds <- hill_pair_dis$q1S %>% #select the distance object based on dissimilarity metric S
  metaMDS(.,trymax = 500, k = 2, verbosity = FALSE) %>%
  vegan:::scores.metaMDS() %>%
  as_tibble(., rownames = "sample")

hill_pair_dis_nmds <- hill_pair_dis_nmds %>%
  left_join(Metadata, by = join_by(sample == SampleID)) %>%
  group_by(Species, Season) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup()
```

### Plot ordination q1


```{r}
NMDS_Plot_q1 <- ggplot(hill_pair_dis_nmds, aes(x = NMDS1, y = NMDS2, 
                                               color = Species, shape = Season)) +
  geom_point(size = 2) +
  geom_segment(aes(x = x_cen, y = y_cen, xend = NMDS1, yend = NMDS2), alpha = 0.2) +
  theme(legend.position = "right",
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  theme_classic() +
  theme(legend.text = element_text(face = "italic")) +
  scale_colour_manual(values = c("#999999","#E69F00","#56B4E9","#009E73","#F0E442"))
print(NMDS_Plot_q1)
#ggsave("../figures/NMDS_Plot_q1.jpeg", width = 8.0, height = 4.0, dpi = 300)
```

### Run PERMANONA
```{r}
set.seed(123)
dist <- hill_pair_dis$q1S
perm_q1 <- adonis2(dist ~ Species*Season+Elevation+SVL+SeqDepth,
                   data = Metadata, permutations = 999)
print(perm_q1)
```



### Run pairwise perMANOVA
```{r}
Metadata$species_season <- paste(Metadata$Species, Metadata$Season, sep = "_")
pw_adonis_q1<-pairwise.adonis(dist,factors=Metadata$species_season)

intrasp_compar <- pw_adonis_q1[grep("^([^_]+)_Dry vs \\1_Rainy", pw_adonis_q1$pairs), ]
print(intrasp_compar)
```

### Plot of intraspecific distances

```{r}
compar_intrasp<- c("Sceloporus grammicus_Dry_vs_Sceloporus grammicus_Rainy", "Sceloporus aeneus_Dry_vs_Sceloporus aeneus_Rainy", "Sceloporus bicanthalis_Dry_vs_Sceloporus bicanthalis_Rainy", "Sceloporus spinosus_Dry_vs_Sceloporus spinosus_Rainy")

dist.q1 <- 1-(hill_pair_dis$q1S) %>% as.matrix()

#convert matrix into df
dist.q1[lower.tri(dist.q1)] <- NA
dist.q1.df <- as.matrix(dist.q1) %>% melt(
  varnames = c("site1", "site2"))%>% drop_na() %>% filter(!value==0)

#Join beta diversity vakues with metadata
dist.q1.meta<- dist.q1.df %>% 
  inner_join(Metadata, by = c("site1"="SampleID")) %>% 
  inner_join(Metadata, by = c("site2"="SampleID"))

#Subset to obtain specific comparisons 
dist.q1.meta.filt <- dist.q1.meta %>% unite(
  "compar_intraspp", c("species_season.x", "species_season.y"), sep="_vs_", remove=F) %>% filter(compar_intraspp %in% compar_intrasp) 

# Plotting

#Setting the order of the species
dist.q1.meta.filt$Species.order<- factor(dist.q1.meta.filt$Species.x,
                                         levels = c("Sceloporus aeneus", "Sceloporus bicanthalis", "Sceloporus grammicus", "Sceloporus spinosus"))

# Kruskal wallis test
set.seed(1234)
kruskal.t.q1 <- dist.q1.meta.filt %>% kruskal_test(value ~ Species.order)
Post_hoc.q1  <- dist.q1.meta.filt %>% dunn_test(value ~ Species.order, p.adjust.method = "bonferroni") 
Post_hoc.q1 <- Post_hoc.q1 %>% 
  add_xy_position(x = "Species.order")

# Boxplot
q1_intrasp<-dist.q1.meta.filt %>% 
  ggboxplot(x = "Species.order", y="value", fill = "Species.order")+
  ylab("Seasonal turnover of frequent ASVs (q=1)")+
  xlab(element_blank())+
  scale_fill_manual(values = c("#999999","#E69F00","#56B4E9","#009E73","#F0E442"))+
      theme_classic() +
   theme(legend.position = "right",
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank()) +
  theme(legend.text = element_text(face = "italic"))+
  stat_pvalue_manual(Post_hoc.q1, 
                     hide.ns = TRUE, 
                     step.increase = 0.2) +
  labs(
    subtitle = get_test_label(kruskal.t.q1, 
                              detailed = TRUE),
    caption = get_pwc_label(Post_hoc.q1))+
  scale_y_continuous(limits = c(0,1))

  
q1_intrasp
#ggsave("../figures/beta_div_intrasp_q1.jpeg", width = 5.0, height = 4.0, dpi = 300)


```      

### Run BETADISPER analysis
```{r}
disp.species = betadisper(dist, Metadata$Species)
anova(disp.species)
permutest(disp.species)
permutest(disp.species, pairwise = TRUE, permutations = 999)
boxplot(disp.species)
```



## Taxonomic beta diversity  at order q2

```{r, include=FALSE}
hill_pair_dis_nmds_q2 <- hill_pair_dis$q2S %>% #select the distance object based on dissimilarity metric S
  metaMDS(.,trymax = 500, k = 2, verbosity = FALSE) %>%
  vegan:::scores.metaMDS() %>%
  as_tibble(., rownames = "sample")

hill_pair_dis_nmds_2 <- hill_pair_dis_nmds_q2 %>%
  left_join(Metadata, by = join_by(sample == SampleID)) %>%
  group_by(Species, Season) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup()
```

```{r, eval=FALSE}
hill_pair_dis_nmds_q2 <- hill_pair_dis$q2S %>% #select the distance object based on dissimilarity metric S
  metaMDS(.,trymax = 500, k = 2, verbosity = FALSE) %>%
  vegan:::scores.metaMDS() %>%
  as_tibble(., rownames = "sample")

hill_pair_dis_nmds_2 <- hill_pair_dis_nmds_q2 %>%
  left_join(Metadata, by = join_by(sample == SampleID)) %>%
  group_by(Species, Season) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup()
```

### Plot ordination q2

```{r}
NMDS_Plot_q2 <- ggplot(hill_pair_dis_nmds_2, aes(x = NMDS1, y = NMDS2, 
                                                  color = Species, shape = Season)) +
  geom_point(size = 2) +
  geom_segment(aes(x = x_cen, y = y_cen, xend = NMDS1, yend = NMDS2), alpha = 0.2) +
  theme(legend.position = "right",
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  theme_classic() +
  theme(legend.text = element_text(face = "italic")) +
  scale_colour_manual(values = c("#999999","#E69F00","#56B4E9","#009E73","#F0E442"))
print(NMDS_Plot_q2)
#ggsave("../figures/NMDS_Plot_q2.jpeg", width = 8.0, height = 4.0, dpi = 300)
```

### Run PERMANOVA

```{r}

set.seed(123)
dist2 <- hill_pair_dis$q2S
perm_q2 <- adonis2(dist2 ~ Species*Season+Elevation+SVL+SeqDepth,
                   data = Metadata, permutations = 999)
print(perm_q2)
```

### Run pairwise perMANOVA
```{r}

pw_adonis_q2<-pairwise.adonis(dist2,factors=Metadata$species_season)
intrasp_compar_q2 <- pw_adonis_q2[grep("^([^_]+)_Dry vs \\1_Rainy", pw_adonis_q2$pairs), ]
print(intrasp_compar_q2)
```

### Plot of intraspecific distances

```{r}
compar_intrasp<- c("Sceloporus grammicus_Dry_vs_Sceloporus grammicus_Rainy", "Sceloporus aeneus_Dry_vs_Sceloporus aeneus_Rainy", "Sceloporus bicanthalis_Dry_vs_Sceloporus bicanthalis_Rainy", "Sceloporus spinosus_Dry_vs_Sceloporus spinosus_Rainy")

dist.q2 <- 1-(hill_pair_dis$q2S) %>% as.matrix()

#convert matrix into df
dist.q2[lower.tri(dist.q2)] <- NA
dist.q2.df <- as.matrix(dist.q2) %>% melt(
  varnames = c("site1", "site2"))%>% drop_na() %>% filter(!value==0)

#Join beta diversity vakues with metadata
dist.q2.meta<- dist.q2.df %>% 
  inner_join(Metadata, by = c("site1"="SampleID")) %>% 
  inner_join(Metadata, by = c("site2"="SampleID"))

#Subset to obtain specific comparisons 
dist.q2.meta.filt <- dist.q2.meta %>% unite(
  "compar_intraspp", c("species_season.x", "species_season.y"), sep="_vs_", remove=F) %>% filter(compar_intraspp %in% compar_intrasp) 

# Plotting

#Setting the order of the species
dist.q2.meta.filt$Species.order<- factor(dist.q2.meta.filt$Species.x,
                                         levels = c("Sceloporus aeneus", "Sceloporus bicanthalis", "Sceloporus grammicus", "Sceloporus spinosus"))

# Kruskal wallis test
set.seed(1234)
kruskal.t.q2 <- dist.q2.meta.filt %>% kruskal_test(value ~ Species.order)
Post_hoc.q2  <- dist.q2.meta.filt %>% dunn_test(value ~ Species.order, p.adjust.method = "bonferroni") 
Post_hoc.q2 <- Post_hoc.q2 %>% 
  add_xy_position(x = "Species.order")

q2_intrasp<-dist.q2.meta.filt %>% 
 ggboxplot(x = "Species.order", y="value", fill = "Species.order")+
  ylab("Seasonal turnover of dominant ASVs (q=2)")+
  xlab(element_blank())+
  scale_fill_manual(values = c("#999999","#E69F00","#56B4E9","#009E73","#F0E442"))+
      theme_classic() +
   theme(legend.position = "right",
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank()) +
  theme(legend.text = element_text(face = "italic"))+
  stat_pvalue_manual(Post_hoc.q2, 
                     hide.ns = TRUE,
                     step.increase = 0.1) +
  labs(
    subtitle = get_test_label(kruskal.t.q2, 
                              detailed = TRUE),
    caption = get_pwc_label(Post_hoc.q2))+
  scale_y_continuous(limits = c(0,1))
  
q2_intrasp
#ggsave("../figures/beta_div_intrasp_q2.jpeg", width = 5.0, height = 4.0, dpi = 300)

```                        
                        
                        
### Run BETADISPER analysis
```{r}
disp.species_q2 = betadisper(dist2, Metadata$Species)
anova(disp.species_q2)
permutest(disp.species_q2)
permutest(disp.species_q2, pairwise = TRUE, permutations = 999)
boxplot(disp.species_q2)
```
