---
title: "Analysis_procrustes"
author: 
- name: Mauricio Hernández
  affiliation: Laboratory of Biotic Interactions, Centro de Investigación en Ciencias
    Biológicas, Universidad Autónoma de Tlaxcala, Tlaxcala, Mexico
  email:  delmer.hernandez@unah.edu.hn
- name: Yendi E. Navarro-Noya
  affiliation: Laboratory of Biotic Interactions, Centro de Investigación en Ciencias
    Biológicas, Universidad Autónoma de Tlaxcala, Tlaxcala, Mexico
  email: yendiebenezer.navarro.n@uatx.mx
date: "`r format(Sys.time(), '%d - %m - %Y')`"
output: 
  html_document:
    theme: flatly
    highlight: tango
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
---

# IV. MANTEL Y PROCRUSTES ANALYSIS

## Load libraries
```{r setup, warning=FALSE, message=FALSE}
library(vegan)
library(ggplot2)
library(cowplot)
library(corrr)
library(tidyverse)
library(ape)
library(hilldiv2)
```

## Loading files

This analysis is based on 68 samples obtained from both datasets (diet-microbiota)
```{r}
metadata <- read.csv("../data/Procrustes_metadata.csv", header = TRUE, check.names = FALSE, 
                     row.names = 1)
diet <- read.csv("../data/Procrustes_Diet.csv", header = TRUE, check.names = FALSE, 
                 row.names = 1) %>% t()
microbiota <- read.csv("../data/Procrustes_Microbiota.csv", header = TRUE, 
                       check.names = FALSE, row.names = 1)
#microbiota <- microbiota %>% filter(rowSums(across(where(is.numeric)))!=0) %>% t()

```


## Calculate distances matrices

```{r}
# Distance matrix of bacterial microbiota
jaccard_mic_1 <- hillpair(data = microbiota, q = 1) # distance matrix at order q1
microbiota_jaccard_1 <- as.dist(jaccard_mic_1$S)
jaccard_mic_2 <- hillpair(data = microbiota, q = 2) # distance matrix at order q1
microbiota_jaccard_2 <- as.dist(jaccard_mic_2$S)

# Distance matrix of diet
diet_jaccard <- vegdist(diet, method = "jaccard")


```


## Mantel test
```{r}
set.seed(123)
# Order q=1
mantel(diet_jaccard, microbiota_jaccard_1, method = "pearson",
       permutations = 999)

#Order q=2
mantel(diet_jaccard, microbiota_jaccard_2, method = "pearson",
       permutations = 999)
```



## Make pcoas 

```{r}

#PcoA of diet
dietPCoA <- as.data.frame(cmdscale(diet_jaccard))
plot(dietPCoA)

# PCoA of microbiota q=1
microbiotaPCoA_1 <- as.data.frame(cmdscale(microbiota_jaccard_1))
plot(microbiotaPCoA_1)

# PCoA of microbiota q=2
microbiotaPCoA_2 <- as.data.frame(cmdscale(microbiota_jaccard_2))
plot(microbiotaPCoA_2)
```



## Procrustes analysis of all species

### Procrustes analysis of all species at order q=1
```{r}

procrust <- procrustes(X = dietPCoA, Y = microbiotaPCoA_1, scale=TRUE,
                       symmetric = TRUE)
pro_test <- protest(dietPCoA, microbiotaPCoA_1, permutations = 9999)
pro_test
plot(pro_test)

eigen <- sqrt(procrust$svd$d)
percent_var <- signif(eigen/sum(eigen), 4)*100

beta_pro <- data.frame(procrust$X)
head(beta_pro)
trans_pro <- data.frame(procrust$Yrot)
head(trans_pro)
beta_pro$UserName <- rownames(beta_pro)
beta_pro$type <- "Diet"
beta_pro$seasons<-metadata[,5]
beta_pro$species<-metadata[,1]
head(beta_pro)

trans_pro$UsarName <- rownames(trans_pro)
trans_pro$type <- "Microbiota"
trans_pro$seasons=metadata[,5]
trans_pro$species=metadata[,1]
head(trans_pro)

colnames(trans_pro) <- colnames(beta_pro)
toplot <- data.frame(rbind(beta_pro,trans_pro))
pval <- signif(pro_test$signif, 1)

#col1 = rgb(250/255,60/255,60/255)
#col4 = rgb(0/255,200/255,200/255)
#col8 = rgb(160/255,0/255,200/255)
#col10 = rgb(0/255,160/255,255/255)
# "#999999","#E69F00","#56B4E9", "#009E73","#F0E442

```



### Graph diet and microbiota

```{r}
procrustes_new <- ggplot(toplot) +
  geom_point(size = 1.8, alpha = 0.75, aes(x = V1, y = V2, color = species, 
                                         shape = type)) + 
  scale_color_manual(values = c("#999999","#E69F00","#56B4E9",
                                "#009E73")) +
  theme_classic() +
  scale_x_continuous(limits = c(-0.1, 0.3)) +
  scale_y_continuous(limits = c(-0.2, 0.2)) +
  geom_line(aes(x= V1, y = V2, group = UserName), 
            col = "darkgrey", alpha = 0.6, 
            size = 0.2) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 10, colour = "black"),
        legend.position = "right",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12, colour = "black"),
        aspect.ratio = 1) +
  theme(legend.text = element_text(face = "italic")) +
  guides(color = guide_legend(ncol = 1)) +
  annotate("text", x = 0.2, y = -0.2, 
           label = paste0("p-value=", pval), size = 4) +
  xlab(paste0("PCoA 1 (",percent_var[1],"%)")) +
  ylab(paste0("PCoA 2 (",percent_var[2],"%)"))
print(procrustes_new)
ggsave("../figures/procrustes_diet_q1.jpeg", width = 8.0, height = 4.0, dpi = 300)
```

### Procrustes analysis of all species at order q=2
```{r}

procrust <- procrustes(X = dietPCoA, Y = microbiotaPCoA_2, scale=TRUE,
                       symmetric = TRUE)
pro_test <- protest(dietPCoA, microbiotaPCoA_2, permutations = 9999)
pro_test
plot(pro_test)

eigen <- sqrt(procrust$svd$d)
percent_var <- signif(eigen/sum(eigen), 4)*100

beta_pro <- data.frame(procrust$X)
head(beta_pro)
trans_pro <- data.frame(procrust$Yrot)
head(trans_pro)
beta_pro$UserName <- rownames(beta_pro)
beta_pro$type <- "Diet"
beta_pro$seasons<-metadata[,5]
beta_pro$species<-metadata[,1]
head(beta_pro)

trans_pro$UsarName <- rownames(trans_pro)
trans_pro$type <- "Microbiota"
trans_pro$seasons=metadata[,5]
trans_pro$species=metadata[,1]
head(trans_pro)

colnames(trans_pro) <- colnames(beta_pro)
toplot <- data.frame(rbind(beta_pro,trans_pro))
pval <- signif(pro_test$signif, 1)

```



### Graph diet and microbiota

```{r}
procrustes_new <- ggplot(toplot) +
  geom_point(size = 1.8, alpha = 0.75, aes(x = V1, y = V2, color = species, 
                                         shape = type)) + 
  scale_color_manual(values = c("#999999","#E69F00","#56B4E9",
                                "#009E73")) +
  theme_classic() +
  scale_x_continuous(limits = c(-0.1, 0.3)) +
  scale_y_continuous(limits = c(-0.2, 0.2)) +
  geom_line(aes(x= V1, y = V2, group = UserName), 
            col = "darkgrey", alpha = 0.6, 
            size = 0.2) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 10, colour = "black"),
        legend.position = "right",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12, colour = "black"),
        aspect.ratio = 1) +
  theme(legend.text = element_text(face = "italic")) +
  guides(color = guide_legend(ncol = 1)) +
  annotate("text", x = 0.2, y = -0.2, 
           label = paste0("p-value=", pval), size = 4) +
  xlab(paste0("PCoA 1 (",percent_var[1],"%)")) +
  ylab(paste0("PCoA 2 (",percent_var[2],"%)"))
print(procrustes_new)
ggsave("../figures/procrustes_diet_q2.jpeg", width = 8.0, height = 4.0, dpi = 300)
```



## Procrustes analysis by lizard species

Procrsutes by lizard species were done at q=1 as similar patterns were observed in the general procrustes at q=2.

### Procrustes of  Sceloporus aeneus

```{r}

#loading data
aeneus_metadata <- subset(metadata, Species_Scel == "Sceloporus aeneus")
diet_aeneus <- read.csv("../data/Proc_Diet_aeneus.csv", header = TRUE, check.names = FALSE, 
                        row.names = 1) %>% t()
microbiota_aeneus <- read.csv("../data/Proc_Mic_aeneus.csv", header = TRUE, 
                              check.names = FALSE, row.names = 1)

# Distance matrix
jaccard_mic_a <- hillpair(data = microbiota_aeneus, q = 1)
microbiota_jaccard_a <- as.dist(jaccard_mic_a$S)

# Absence/presence data frame - Jaccard distance
diet_jaccard_a <- vegdist(diet_aeneus, method = "jaccard")
diet_jaccard_a <- as.dist(diet_jaccard_a)

# Mantel test

set.seed(123)
mantel(diet_jaccard_a, microbiota_jaccard_a, method = "pearson",
       permutations = 999)


# Make pcoas

dietPCoA_a <- as.data.frame(cmdscale(diet_jaccard_a))
plot(dietPCoA_a)
microbiotaPCoA_a <- as.data.frame(cmdscale(microbiota_jaccard_a))
plot(microbiotaPCoA_a)


# Procrustes analysis

procrust_a <- procrustes(X = dietPCoA_a, Y = microbiotaPCoA_a, scale=TRUE,
                         symmetric = TRUE)
pro_test_a <- protest(dietPCoA_a, microbiotaPCoA_a, permutations = 9999)
pro_test_a
plot(pro_test_a)

eigen_a <- sqrt(procrust_a$svd$d)
percent_var_a <- signif(eigen_a/sum(eigen_a), 4)*100

beta_pro_a <- data.frame(procrust_a$X)
head(beta_pro_a)
trans_pro_a <- data.frame(procrust_a$Yrot)
head(trans_pro_a)
beta_pro_a$UserName <- rownames(beta_pro_a)
beta_pro_a$type <- "Diet"
beta_pro_a$seasons <- aeneus_metadata[,5]
beta_pro_a$species <- aeneus_metadata[,1]
head(beta_pro_a)

trans_pro_a$UsarName <- rownames(trans_pro_a)
trans_pro_a$type <- "Microbiota"
trans_pro_a$seasons <- aeneus_metadata[,5]
trans_pro_a$species <- aeneus_metadata[,1]
head(trans_pro_a)

colnames(trans_pro_a) <- colnames(beta_pro_a)
toplot_a <- data.frame(rbind(beta_pro_a, trans_pro_a))
pval_a <- signif(pro_test_a$signif, 1)


# Graph diet and microbiota

procreustes_aeneus_new <- ggplot(toplot_a) +
  geom_point(size = 2, alpha = 0.75, aes(x = V1, y = V2, color = species, 
                                         shape = type)) + 
  scale_color_manual(values = c("#999999")) +
  theme_classic() +
  scale_x_continuous(limits = c(-0.2, 0.4)) +
  scale_y_continuous(limits = c(-0.2, 0.4)) +
  geom_line(aes(x = V1, y = V2, group = UserName), 
            col = "darkgrey", alpha = 0.6, 
            size = 0.2) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 10, colour = "black"),
        legend.position = "right",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12, colour = "black"),
        aspect.ratio = 1) +
  theme(legend.text = element_text(face = "italic")) +
  guides(color = guide_legend(ncol = 1)) +
  #annotate("text", x = 0.07, y = -0.13, label = paste0("p-value=", pval_a), 
  #         size = 4) +
  xlab(paste0("PCoA 1 (",percent_var_a[1],"%)")) +
  ylab(paste0("PCoA 2 (",percent_var_a[2],"%)"))
print(procreustes_aeneus_new)
ggsave("../figures/procreustes_aeneus_new.jpeg", width = 8.0, height = 4.0, dpi = 300)
```


### Procrustes of Sceloporus bicanthalis 
```{r}

#Loading data
bica_metadata <- subset(metadata, Species_Scel == "Sceloporus bicanthalis")
diet_bica <- read.csv("../data/Proc_Diet_bica.csv", header = TRUE, check.names = FALSE, 
                      row.names = 1) %>% t()
microbiota_bica <- read.csv("../data/Proc_Mic_bica.csv", header = TRUE, 
                            check.names = FALSE, row.names = 1)

# Distance matrix
jaccard_mic_b <- hillpair(data = microbiota_bica, q = 1)
microbiota_jaccard_b <- as.dist(jaccard_mic_b$S)

# Absence/presence data frame - Jaccard distance
diet_jaccard_b <- vegdist(diet_bica, method = "jaccard")
diet_jaccard_b <- as.dist(diet_jaccard_b)


# Mantel test
set.seed(123)
mantel(diet_jaccard_b, microbiota_jaccard_b, method = "pearson",
       permutations = 999)

# Make pcoas
dietPCoA_b <- as.data.frame(cmdscale(diet_jaccard_b))
plot(dietPCoA_b)

microbiotaPCoA_b <- as.data.frame(cmdscale(microbiota_jaccard_b))
plot(microbiotaPCoA_b)

# Procrustes analysis
procrust_b <- procrustes(X = dietPCoA_b, Y = microbiotaPCoA_b, scale=TRUE,
                         symmetric = TRUE)
pro_test_b <- protest(dietPCoA_b, microbiotaPCoA_b, permutations = 9999)
pro_test_b
plot(pro_test_b)

eigen_b <- sqrt(procrust_b$svd$d)
percent_var_b <- signif(eigen_b/sum(eigen_b), 4)*100

beta_pro_b <- data.frame(procrust_b$X)
head(beta_pro_b)
trans_pro_b <- data.frame(procrust_b$Yrot)
head(trans_pro_b)
beta_pro_b$UserName <- rownames(beta_pro_b)
beta_pro_b$type <- "Diet"
beta_pro_b$seasons <- bica_metadata[,5]
beta_pro_b$species <- bica_metadata[,1]
head(beta_pro_b)

trans_pro_b$UsarName <- rownames(trans_pro_b)
trans_pro_b$type <- "Microbiota"
trans_pro_b$seasons <- bica_metadata[,5]
trans_pro_b$species <- bica_metadata[,1]
head(trans_pro_b)

colnames(trans_pro_b) <- colnames(beta_pro_b)
toplot_b <- data.frame(rbind(beta_pro_b, trans_pro_b))
pval_b <- signif(pro_test_b$signif, 1)

# graph diet and microbiota
procreustes_bica_new <- ggplot(toplot_b) +
  geom_point(size = 2, alpha = 0.75, aes(x = V1, y = V2, color = species, 
                                         shape = type)) + 
  scale_color_manual(values = c("#E69F00")) +
  theme_classic() +
  #scale_x_continuous(limits = c(-0.25, 0.4)) +
  #scale_y_continuous(limits = c(-0.2, 0.25)) +
  geom_line(aes(x = V1, y = V2, group = UserName), 
            col = "darkgrey", alpha = 0.6, 
            size = 0.2) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 10, colour = "black"),
        legend.position = "right",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12, colour = "black"),
        aspect.ratio = 1) +
  theme(legend.text = element_text(face = "italic")) +
  guides(color = guide_legend(ncol = 1)) +
  #annotate("text", x = 0.07, y = -0.28, label = paste0("p-value=", pval_b), 
  #        size = 4) +
  xlab(paste0("PCoA 1 (",percent_var_b[1],"%)")) +
  ylab(paste0("PCoA 2 (",percent_var_b[2],"%)"))
print(procreustes_bica_new)
ggsave("../figures/procreustes_bica_new.jpeg", width = 8.0, height = 4.0, dpi = 300)

```



### Procrustes: Sceloporus grammicus

```{r}

# Loading data

gram_metadata <- subset(metadata, Species_Scel == "Sceloporus grammicus")
diet_gram <- read.csv("../data/Proc_Diet_gram.csv", header = TRUE, check.names = FALSE, 
                      row.names = 1) %>% t()
microbiota_gram <- read.csv("../data/Proc_Mic_gram.csv", header = TRUE, 
                            check.names = FALSE, row.names = 1)

# Distance matrix
jaccard_mic_g <- hillpair(data = microbiota_gram, q = 1)
microbiota_jaccard_g <- as.dist(jaccard_mic_g$S)

# Absence/presence data frame - Jaccard distance
diet_jaccard_g <- vegdist(diet_gram, method = "jaccard")
diet_jaccard_g <- as.dist(diet_jaccard_g)

# Mantel test
set.seed(123)
mantel(diet_jaccard_g, microbiota_jaccard_g, method = "pearson",
       permutations = 999)

# Make pcoas
dietPCoA_g <- as.data.frame(cmdscale(diet_jaccard_g))
plot(dietPCoA_g)

microbiotaPCoA_g <- as.data.frame(cmdscale(microbiota_jaccard_g))
plot(microbiotaPCoA_g)

# Procrustes analysis
procrust_g <- procrustes(X = dietPCoA_g, Y = microbiotaPCoA_g, scale=TRUE,
                         symmetric = TRUE)
pro_test_g <- protest(dietPCoA_g, microbiotaPCoA_g, permutations = 9999)
pro_test_g
plot(pro_test_g)

eigen_g <- sqrt(procrust_g$svd$d)
percent_var_g <- signif(eigen_g/sum(eigen_g), 4)*100

beta_pro_g <- data.frame(procrust_g$X)
head(beta_pro_g)
trans_pro_g <- data.frame(procrust_g$Yrot)
head(trans_pro_g)
beta_pro_g$UserName <- rownames(beta_pro_g)
beta_pro_g$type <- "Diet"
beta_pro_g$seasons <- gram_metadata[,5]
beta_pro_g$species <- gram_metadata[,1]
head(beta_pro_g)

trans_pro_g$UsarName <- rownames(trans_pro_g)
trans_pro_g$type <- "Microbiota"
trans_pro_g$seasons <- gram_metadata[,5]
trans_pro_g$species <- gram_metadata[,1]
head(trans_pro_g)

colnames(trans_pro_g) <- colnames(beta_pro_g)
toplot_g <- data.frame(rbind(beta_pro_g, trans_pro_g))
pval_g <- signif(pro_test_g$signif, 1)

# Graph diet and microbiota
procreustes_gram_new <- ggplot(toplot_g) +
  geom_point(size = 2, alpha = 0.75, aes(x = V1, y = V2, color = species, 
                                         shape = type)) + 
  scale_color_manual(values = c("#56B4E9")) +
  theme_classic() +
  #scale_x_continuous(limits = c(-0.25, 0.4)) +
  #scale_y_continuous(limits = c(-0.2, 0.25)) +
  geom_line(aes(x = V1, y = V2, group = UserName), 
            col = "darkgrey", alpha = 0.6, 
            size = 0.2) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 10, colour = "black"),
        legend.position = "right",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12, colour = "black"),
        aspect.ratio = 1) +
  theme(legend.text = element_text(face = "italic")) +
  guides(color = guide_legend(ncol = 1)) +
  #annotate("text", x = 0.07, y = -0.28, label = paste0("p-value=", pval_b), 
  #        size = 4) +
  xlab(paste0("PCoA 1 (",percent_var_g[1],"%)")) +
  ylab(paste0("PCoA 2 (",percent_var_g[2],"%)"))
print(procreustes_gram_new)
ggsave("../figures/procreustes_gram_new.jpeg", width = 8.0, height = 4.0, dpi = 300)

```



### Procrustes: Sceloporus spinosus #
```{r}


spi_metadata <- subset(metadata, Species_Scel == "Sceloporus spinosus")
diet_spi <- read.csv("../data/Proc_Diet_spi.csv", header = TRUE, check.names = FALSE, 
                     row.names = 1) %>% t()
microbiota_spi <- read.csv("../data/Proc_Mic_spi.csv", header = TRUE, 
                           check.names = FALSE, row.names = 1)

# Distance matrix
jaccard_mic_s <- hillpair(data = microbiota_spi, q = 1)
microbiota_jaccard_s <- as.dist(jaccard_mic_s$S)

# Absence/presence data frame - Jaccard distance
diet_jaccard_s <- vegdist(diet_spi, method = "jaccard")
diet_jaccard_s <- as.dist(diet_jaccard_s)


# Mantel test 
set.seed(123)
mantel(diet_jaccard_s, microbiota_jaccard_s, method = "pearson",
       permutations = 999)

# Make pcoas
dietPCoA_s <- as.data.frame(cmdscale(diet_jaccard_s))
plot(dietPCoA_s)

microbiotaPCoA_s <- as.data.frame(cmdscale(microbiota_jaccard_s))
plot(microbiotaPCoA_s)

# Procrustes analysis
procrust_s <- procrustes(X = dietPCoA_s, Y = microbiotaPCoA_s, scale=TRUE,
                         symmetric = TRUE)
pro_test_s <- protest(dietPCoA_s, microbiotaPCoA_s, permutations = 9999)
pro_test_s
plot(pro_test_s)

eigen_s <- sqrt(procrust_s$svd$d)
percent_var_s <- signif(eigen_s/sum(eigen_s), 4)*100

beta_pro_s <- data.frame(procrust_s$X)
head(beta_pro_s)
trans_pro_s <- data.frame(procrust_s$Yrot)
head(trans_pro_s)
beta_pro_s$UserName <- rownames(beta_pro_s)
beta_pro_s$type <- "Diet"
beta_pro_s$seasons <- spi_metadata[,5]
beta_pro_s$species <- spi_metadata[,1]
head(beta_pro_s)

trans_pro_s$UsarName <- rownames(trans_pro_s)
trans_pro_s$type <- "Microbiota"
trans_pro_s$seasons <- spi_metadata[,5]
trans_pro_s$species <- spi_metadata[,1]
head(trans_pro_s)

colnames(trans_pro_s) <- colnames(beta_pro_s)
toplot_s <- data.frame(rbind(beta_pro_s, trans_pro_s))
pval_s <- signif(pro_test_s$signif, 1)

# Graph diet and microbiota
procreustes_spi_new <- ggplot(toplot_s) +
  geom_point(size = 2, alpha = 0.75, aes(x = V1, y = V2, color = species, 
                                         shape = type)) + 
  scale_color_manual(values = c("#40A175")) +
  theme_classic() +
  #scale_x_continuous(limits = c(-0.25, 0.4)) +
  #scale_y_continuous(limits = c(-0.2, 0.25)) +
  geom_line(aes(x = V1, y = V2, group = UserName), 
            col = "darkgrey", alpha = 0.6, 
            size = 0.2) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 10, colour = "black"),
        legend.position = "right",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12, colour = "black"),
        aspect.ratio = 1) +
  theme(legend.text = element_text(face = "italic")) +
  guides(color = guide_legend(ncol = 1)) +
  #annotate("text", x = 0.07, y = -0.28, label = paste0("p-value=", pval_s), 
  #        size = 4) +
  xlab(paste0("PCoA 1 (",percent_var_s[1],"%)")) +
  ylab(paste0("PCoA 2 (",percent_var_s[2],"%)"))
print(procreustes_spi_new)
ggsave("../figures/procreustes_spi_new.jpeg", width = 8.0, height = 4.0, dpi = 300)
```

## Procustes and mantel analysis considering phylogenetic comosition of diet

```{r, eval=FALSE}
tree.diet <- read.tree("../data/tree.nwk")
comm<-(diet)
IDs= colnames(comm)   # para obtener los nombres de las sp
tree.diet= ape::rtree(n=ncol(diet), tip.label = paste0(IDs))  # tener el árbol enraizado, con los nombres de los sitios y los nombres de las sp.


hill_phylo_parti_pairwise(comm,tree.diet, q = 0, show_warning = F) 

```
  
```{r, eval=FALSE}  
# Loading phylogenetic tree
phy_tree <- read_tree("../data/tree.nwk")

# Creating phyloseq object
SAM <- sample_data(metadata)
OTU <- otu_table(diet, taxa_are_rows=F)
phylo_physeq = phyloseq(OTU, SAM, phy_tree)

#Calculating phylogenetic distance
diet_dist_phy = distance(phylo_physeq, method = "uunifrac") %>% as.dist

```

```{r, eval=FALSE}
set.seed(123)
# Order q=1
mantel(diet_dist_phy, microbiota_jaccard_1, method = "pearson",
       permutations = 999)

#Order q=2
mantel(diet_dist_phy, microbiota_jaccard_2, method = "pearson",
       permutations = 999)
```

